---
title: "Arm length vs. Height"
author: "I. N 7184"
date: "September 26, 2020"
output: html_document
---

```{r, include = FALSE}
library(tidyverse)
set.seed(7184) # set the seed
r<-read.csv("a20a1data.csv") # read the csv file
data<-r[sample(nrow(r), 10), ] # choose 10 random rows from the csv file
attach(data) # use column names as variables in this session
```

## I. Introduction 

##### Using words, give a title to summarise what your assignment #1 is about.
##### This title should also be the title of your RMarkdown file.

##### Describe how your sample data was collected.

In class, we were put into several breakout rooms, where we discussed the
lengths of our forearms and heights. Unfortunately, I wanted to wait for the
data file instead so, I did not write down these values and will be using the
a20a1data.csv file. As per the instructions, I have set the sedd to be the last
four digits of my student number (7184) and randomly sampled 10 data points.

##### Specify your explanatory variable and response variable and, give a brief
##### explanation of your choice

The response, or the dependent variable is the arm length, while the
explanatory, or independent variable is the height. This is because arm length
should be based on the height of a person; depending on how tall a person is,
we can assume, or try to guess, their forearm length.

## II. Exploratory Data Analysis

##### Draw at most 3 plots to visually describe your data. Is your response
##### variable approximately Normal?

```{r, echo=FALSE}
# summary of the data
summary(data)
# creating the SLR
lmod=lm(forearm~height)
# plot the data
ggplot(data, aes(x=height,y=forearm))+geom_point()+geom_smooth(method=lm)+ 
  labs(title="Scatterplot of Forearm to Height Ratio")
# qq plot to show normality
qqnorm(data$forearm, main="Normal QQ Plot for Forearm in cm", cex=0.5)
qqline(data$forearm)
```

Yes, the response variable, ignoring one out-lier, seems to be fairly normal
and gathered around the center of the qq normal plot, which dictates that the
response variable is likely Normally distributed.

##### Numerically describe the centre, spread and any unusual points of your
##### variables/data.

The center, or median, seems to be at 23.25cm for Forearm length while it is
around 166.0cm for Height. The spread, about the 1st and 3rd quartile, seems to
be between 21.25cm-27.75cm for Forearm length and 162.5cm-173.8cm for Height. 
There are a few unusual points, one being (177, 40)
and another being (157, 16). I refuse to believe someone has a forearm length of
40 cm, so there seems to be quite some error in some of the student inputs for
the survey. 

## III. Methods and Model

##### Fit and describe a linear regression model between forearm length and
##### height.

```{r, echo=FALSE}
# creating the SLR
lmod=lm(forearm~height)
# show the summary of the line
summary(lmod)
```

The error seems to be quite large, since some of the points (notably (177, 40))
are quite far away from the SLR model.

##### Are the regression parameters zero?

No, the Y intercept variable is not zero, as seen in the summary window. The 
intercept parameter is -77.2078. The slope parameter is 0.6126.

##### Interpret the estimates of the regression parameters

Because the Y intercept is below zero, nor is it close to zero, it is reasonable
to say that perhaps the relation between forearm and height is not linear. Since
the value of the y-intercept is very negative (estimated to be around -77.2),
one could assume that if the relation were indeed linear, then either the data
provided was very incorrect, or there were out-lier people. Also, the p-value is rounded to 0.0434 as we can see on the summary page.

## IV. Discussions and Limitations

##### Identify a potential lurking variable. Describe at least one other issue
##### or limitation of your fit. You may include at most one plot to support
##### your answer here.

As we can see above, there are obviously some outliers. This time I want to look
at the lower outlier, around (157, 16). Had there been an injury or disability
regarding the forearm of an individual, their data would be outliers, since
their forearm may be much shorter or longer than normal. Another issue with
the data given is that perhaps the instructions were unclear and their data was
input as inches instead of cm, thus making that data look incorrect when using
a different unit of measurement.

##### Identify another pair of variables to explore a simple linear regression
##### model. Specify the response variable.

Another possible pair of variables to explore a simple linear regression model 
would be the relationship between height and status. Under the lens of one's
status, whether it be rich or poor or in a first/third world country, we can
look at how much one's height is affected. This would have potential since 
the more affluent have greater access to better nutrients. The response variable
in this case is height while the explanatory variable would be their status.